---
title: "Fraktionierung auf Autorenebene"
author: "Stephan Stahlschmidt and Marion Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
# output: github_document
bibliography: MeineBibliothek.bib
csl: emerald-harvard.csl
output:
  # html_document: default
  # pdf_document: default
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
# setting up Oracle connection to local DB
library("ROracle")
library("knitr")
library("viridis")
library("pool")

# source("credentials_local.R")
source("credentials.R")

# establish sql connection pool
sql_pool <- dbPool(
  drv = Oracle(),
  # dbname = "(DESCRIPTION=(ADDRESS=(PROTOCOL=tcp)(HOST=srv10-oracle)(PORT=1521))(CONNECT_DATA=(SID=ifqdata)))",# local 
  dbname = "(DESCRIPTION=(ADDRESS=(PROTOCOL=tcp)(HOST=127.0.0.1)(PORT=6025))(CONNECT_DATA=(SID=bibliodb01)))",# FIZ server
  username = sql_uid,
  password = sql_pw
)
# checking out one sql connection for sql chunks
conn <- poolCheckout(sql_pool)
```

# Motivation

@waltman_field-normalized_2015 demonstrate that whole counting is incompatible with field normalization. They propose fractional counting on the level of authors implemented by fractional counting on address strings.

In any such approach two issues arise:

1. Several authors might be affiliated to the same organisation
2. Single authors might be affiliated with several organisations

Example:
```{sql, beispiel, connection=conn}
SELECT fk_authors, fk_institutions, type, role, organization1, countrycode
FROM wos_b_2018.items it
LEFT JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items = iai.fk_items
LEFT JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
WHERE ut_eid = '000262133500009'
```

# Potential Solution

Fractional counting based on author attribution and subsequent aggregation to the levle of interest (here: countries) might be computed as follows:

```{sql, loesung, connection=conn}
SELECT fk_items, countrycode, SUM(orga_frak) AS cntry_frac -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution of organsation on author level
      SELECT DISTINCT fk_items, organization1, countrycode, fk_authors,
        (1/(COUNT (DISTINCT fk_authors) OVER (PARTITION BY fk_items))) / 
        (COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items, fk_authors)) AS orga_share
      FROM wos_b_2018.items it
      JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items = iai.fk_items
      JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE ut_eid = '000262133500009'
        AND role = 'author'
        AND type = 'RS'
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode
```

The *KB* versions of *Web of Science* and *Scopus* differ in their usage of *fk_authors*, *type* and *role* (Why?). WoS holds the same *fk_authors* as a corresponding author (*RP*) and as an ordinary author (*RS*), while SCP assigns two different *fk_authors* for the same author which vary in their type (*RP* or *RS*). In any case entries for corresponding authors have to excluded via the *RP* type. Furthermore in the WoS authors have to be separated from other roles via *role = 'author'*.

# Data Quality

Necessary information in variables like *Organization1*, *role* and *type* might not always be available and especially before 2008 the link between authors (*fk_authors*) and their affiliated organisation (*fk_institutions*) was most often only recorded for the corresponding author.

```{r, echo=FALSE, cach=FALSE}
# returning sql connection to pool
poolReturn(conn)
```

```{r, echo=FALSE}
# Article und Review pro Jahr
papersByYear <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT pk_items)
  FROM wos_b_2018.items it
  WHERE pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# inner vs left join between items and items_authors_institutions
outerInneDiff <- dbGetQuery(sql_pool, strwrap(paste0("
SELECT tmp1.pubyear, (tmp2.count_outer - tmp1.count_inner) AS pub_diff
  FROM(
    SELECT pubyear, COUNT(DISTINCT iai.fk_items) count_inner
    FROM wos_b_2018.items it
    JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items=iai.fk_items
    WHERE pubtype = 'Journal'
      AND doctype IN ('Article', 'Review')
      AND pubyear BETWEEN 2007 AND 2017
    GROUP BY pubyear) tmp1
  JOIN(
    SELECT pubyear, COUNT(DISTINCT iai.fk_items) count_outer
    FROM wos_b_2018.items it
    LEFT JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items=iai.fk_items
    WHERE pubtype = 'Journal'
      AND doctype IN ('Article', 'Review')
      AND pubyear BETWEEN 2007 AND 2017
    GROUP BY pubyear) tmp2 ON tmp1.pubyear = tmp2.pubyear
  ORDER BY tmp1.pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  ) 
```

```{r, echo=FALSE}
# Article und Review mit fk_authors als NULL
paperNullAuthor <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM wos_b_2018.items it
  JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE fk_authors IS NULL
    AND pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
    AND (type = 'RS' OR type IS NULL)
    AND (role = 'author' OR role IS NULL)
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# Article und Review mit fk_institutions als NULL
paperNullInstitution <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM wos_b_2018.items it
  JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE fk_institutions IS NULL
    AND pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
    AND (type = 'RS' OR type IS NULL)
    AND (role = 'author' OR role IS NULL)
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# article or review with no information on type
paperNullType <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM wos_b_2018.items it
  JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE type IS NULL
    AND pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
if(dim(paperNullType)[1]==0){
  paperNullType <- as.data.frame(matrix(
    c(2007:2017,rep(0,11)),
    nrow=11,
    ncol=2
  ))
}
```

```{r, echo=FALSE}
# article or review with no information on role
paperNullRole <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM wos_b_2018.items it
  JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE role IS NULL
    AND pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# article or review with no information on organization1
paperNullorganization1 <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM wos_b_2018.items it
  JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
  WHERE organization1 IS NULL
    AND pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# all potential missing information jointly
paperOverall <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(pk_items)
  FROM(
  SELECT DISTINCT pubyear, pk_items
  FROM wos_b_2018.items it
  JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items = iai.fk_items
  WHERE pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
    AND ((fk_institutions IS NULL
          AND (type = 'RS' OR type IS NULL) AND (role = 'author' OR role IS NULL)) OR
        (fk_authors IS NULL
          AND (type = 'RS' OR type IS NULL) AND (role = 'author' OR role IS NULL)) OR
        (type IS NULL) OR
        (role IS NULL))
  UNION
  SELECT DISTINCT pubyear,  pk_items
  FROM wos_b_2018.items it
  JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
  WHERE organization1 IS NULL
    AND pubtype = 'Journal'
    AND doctype IN ('Article', 'Review')
    AND pubyear BETWEEN 2007 AND 2017
  )
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

Percentages of pk_items with partially missing information:

```{r, tab_friction, echo=FALSE, cache=FALSE}
# print(outerInneDiff[,2])
# print(paperOverall[,2]/papersByYear[,2])
# print((paperOverall[,2] + outerInneDiff[,2])/papersByYear[,2])
tab_friction <- as.matrix(cbind(outerInneDiff[,2]/papersByYear[,2],
                                paperNullType[,2]/papersByYear[,2],
                                paperNullRole[,2]/papersByYear[,2],
                                paperNullInstitution[,2]/papersByYear[,2],
                                paperNullAuthor[,2]/papersByYear[,2],
                                paperNullorganization1[,2]/papersByYear[,2],
                                (paperOverall[,2] + outerInneDiff[,2])/papersByYear[,2])*100)
colnames(tab_friction) <- c("only pk_item ",
                            "no type",
                            "no role",
                            "no affiliation",
                            "no author",
                            "No Organization1",
                            "total")
rownames(tab_friction) <- as.character(2007:2017)
kable(tab_friction,
      digits = 1
      )
matplot(2007:2017, tab_friction,
     xlim=c(2007,2017),
     ylim=c(0,100),
     lwd=2,
     lty = 1,
     bty="n",
     xlab="",
     ylab="%",
     type="l",
     col = rev(viridis(7)))
legend("topright", legend = rev(colnames(tab_friction)), fill=viridis(7), bty = "n")
```


<!-- ```{r tab_dif_author_counting, echo = FALSE, eval = FALSE} -->
<!-- # Differenz zwischen author_cnt und Zählung der fk_authors -->
<!-- dif_author_counting <- dbGetQuery(sql_pool, strwrap(paste0(" -->
<!--   SELECT pubyear, SUM(diff) -->
<!--   FROM( -->
<!--     SELECT DISTINCT fk_items, pubyear, -->
<!--       CASE  WHEN author_cnt = (COUNT(DISTINCT fk_authors) OVER (PARTITION BY fk_items)) -->
<!--               THEN 0 -->
<!--             ELSE 1 -->
<!--       END diff -->
<!--     FROM wos_b_2018.items it -->
<!--     JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items -->
<!--     WHERE pubtype = 'Journal' -->
<!--       AND doctype IN ('Article', 'Review') -->
<!--       AND pubyear BETWEEN 2007 AND 2017 -->
<!--       AND type = 'RS' -->
<!--       AND role = 'author' -->
<!--   ) -->
<!--   GROUP BY pubyear -->
<!--   ORDER BY pubyear ASC -->
<!--   "), -->
<!--   width = 100000, simplify = TRUE) -->
<!--   ) -->
<!-- ``` -->

<!-- Prozentualer Anteil von pk_items mit Unterschieden in author_cnt und Zählung der zugehörigen fk_authors: -->

<!-- ```{r tab_diff, echo = FALSE, eval = FALSE} -->
<!-- tmp <- as.matrix(dif_author_counting[,2]/papersByYear[,2] * 100) -->
<!-- rownames(tmp) <- as.character(2007:2017) -->
<!-- kable(tmp, -->
<!--       digits = 1, -->
<!--       row.names = TRUE, -->
<!--       col.names=c("# Autoren != author_cnt")) -->
<!-- matplot(2007:2017, tmp, -->
<!--      xlim=c(2007,2017), -->
<!--      ylim=c(0,100), -->
<!--      lwd=2, -->
<!--      lty=1, -->
<!--      bty="n", -->
<!--      xlab="", -->
<!--      ylab="%", -->
<!--      type="l", -->
<!--      col=viridis(2)) -->
<!-- # legend("topright", -->
<!-- #        legend = c("# Autoren != author_cnt", "# Affiliationen != inst_cnt"), -->
<!-- #        fill=viridis(2), -->
<!-- #        bty = "n") -->
<!-- ``` -->

# Implementation

Due to large shares of items with incomplete information we propose a hybrid spproach. We (1) fractionalize on the author level (and subsequently aggregate to the country level) for every pk_item with complete information in the *items_authors_institutions* table and (2) apply fractionalization on the organisation level (and therefore ignore any authors) via *organization1* for all other pk_items. Hence we follow @waltman_field-normalized_2015 but account for data quality issues at the same time.

In detail we compute a table which holds country specific weights for fractional counting. Lower aggregates levels like organisations might also be computed.

```sql
DROP TABLE items_iai_complete;
DROP TABLE wosb2018_frc_cntrylvl;

/*
Create temporaray table as an indicator of items with complete iai record:
*/
CREATE GLOBAL TEMPORARY TABLE items_iai_complete (
    fk_items NUMBER
)
ON COMMIT DELETE ROWS;

/*
create empty table defining fractional weight for every paper by country
causes implicit commit
*/
CREATE TABLE wosb2018_frc_cntrylvl (
    fk_items NUMBER,
    countrycode VARCHAR2(10 CHAR),
    frac_share NUMBER(3,2),
    ind_complete NUMBER(1,0)
);

/*
populate temporary table
*/
INSERT INTO items_iai_complete (
SELECT DISTINCT tmp.fk_items -- subset of items with complete iai information
FROM( -- counts numbers of entries in iai for every item
    SELECT fk_items, COUNT (*) AS entry_cnt
    FROM wos_b_2018.items it
    LEFT JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
    LEFT JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
    WHERE (role IS NULL OR role = 'author') -- we only care for authors and missing information, other roles are neglected
        AND pubtype = 'Journal'
        AND doctype IN ('Article', 'Review')
        AND pubyear BETWEEN 2007 AND 2017
    GROUP BY fk_items
) tmp
JOIN( -- counts numbers of COMPLETE entries in iai for every item
    SELECT fk_items, COUNT (*) AS entry_cnt
    FROM wos_b_2018.items it
    LEFT JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
    LEFT JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
    WHERE pubtype = 'Journal'
        AND doctype IN ('Article', 'Review')
        AND pubyear BETWEEN 2007 AND 2017
        AND fk_institutions IS NOT NULL
        AND fk_authors IS NOT NULL
        AND role = 'author'
        AND type IS NOT NULL
    GROUP BY fk_items
) tmp2 ON tmp.fk_items = tmp2.fk_items
AND tmp.entry_cnt = tmp2.entry_cnt); -- compare both numbers

/*
include items with COMPLETE iai information via fractional counting on author level:
*/
INSERT INTO wosb2018_frc_cntrylvl
SELECT fk_items, countrycode, SUM(orga_frak) AS frac_share, 1 -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution of organsation on author level
      SELECT DISTINCT fk_items, ORGANIZATION1, countrycode, fk_authors,
        (1/(COUNT (DISTINCT fk_authors) OVER (PARTITION BY fk_items)))
        /(COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items, fk_authors)) AS orga_share
      FROM wos_b_2018.items it
      JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
      JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE pubtype = 'Journal'
        AND doctype IN ('Article', 'Review')
        AND pubyear BETWEEN 2007 AND 2017
        AND role = 'author'
        AND type = 'RS'
        AND pk_items IN (SELECT fk_items FROM items_iai_complete)
        AND organization1 IS NOT NULL
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode, 1;

/*
include items with INCOMPLETE iai information via fractional counting on organization level:
*/
INSERT INTO wosb2018_frc_cntrylvl
SELECT fk_items, countrycode, SUM(orga_frak) AS frac_share, 0 -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution on organisation level
      SELECT DISTINCT fk_items, ORGANIZATION1, countrycode,
        (1/(COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items))) AS orga_share
      FROM wos_b_2018.items it
      JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
      JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE pubtype = 'Journal'
        AND doctype IN ('Article', 'Review')
        AND pubyear BETWEEN 2007 AND 2017
        AND (role = 'author' OR role IS NULL)
        AND type = 'RS'
        AND pk_items NOT IN (SELECT fk_items FROM items_iai_complete)
        AND organization1 IS NOT NULL
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode, 0;

CREATE INDEX wosb2018_frc_cntrylvl_ix ON wosb2018_frc_cntrylvl (fk_items, countrycode);

COMMIT;
```

# Caveats

* Whenever an authors holds several affiliations of which two or more actually describe the same organisation, but have not been cleaned in *organization1*, we assign erroneous weights. This issue might be less severe on a country level analysis and could ultimately only be resolved via a worldwide institution harmonization.
* Affiliations without an *organzation1* are omitted, e.g. private addresses (1,2% in wos_b_2018)
* anonymous authors (*role = 'anon'*) are omitted
* Any other?

# Comparison

As an alternative implementation we also compute weights based on fractionation on the *organization1* level:

```sql

/*
create empty table defining fractional weight for every paper by country
based on fractionation on organization1 level
*/
CREATE TABLE wosb2018_frc_cntrylvl_org1 (
    fk_items NUMBER,
    countrycode VARCHAR2(10 CHAR),
    frac_share NUMBER(3,2)
);

/*
include items with INCOMPLETE iai information via fractional counting on organization level:
*/
INSERT INTO wosb2018_frc_cntrylvl_org1
SELECT fk_items, countrycode, SUM(orga_frak) AS frac_share -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution on organisation level
      SELECT DISTINCT fk_items, ORGANIZATION1, countrycode,
        (1/(COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items))) AS orga_share
      FROM wos_b_2018.items it
      JOIN wos_b_2018.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
      JOIN wos_b_2018.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE pubtype = 'Journal'
        AND doctype IN ('Article', 'Review')
        AND pubyear BETWEEN 2007 AND 2017
        AND (role = 'author' OR role IS NULL)
        AND type = 'RS'
        AND organization1 IS NOT NULL
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode;

CREATE INDEX wosb2018_frc_cntrylvl_org1_ix ON wosb2018_frc_cntrylvl_org1 (fk_items, countrycode);

COMMIT;
```

Computing fractional counts of publications on national level for 2017:

```{r, testing1, echo=FALSE, cache=TRUE}
# hybrid approach
pubs_cntry_whole_wos_frac <- dbGetQuery(sql_pool, strwrap(paste0("
    SELECT country, pubyear, SUM(frac_share) pub_cnt_frak
    FROM(
      SELECT DISTINCT it.pubyear, inst.countrycode country, it.pk_items, frak.frac_share
      FROM wos_b_2018.items it
      JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items=iai.fk_items
        AND (iai.role = 'author' OR iai.role IS NULL)
      JOIN wos_b_2018.institutions inst ON inst.pk_institutions=iai.fk_institutions
      JOIN wosb2018_frc_cntrylvl frak ON frak.fk_items = it.pk_items
        AND frak.countrycode = inst.countrycode
      WHERE it.pubyear = 2017
        AND it.doctype IN ('Article','Review')
        AND it.pubtype = 'Journal'
        AND inst.countrycode IN ('AUT', 'BEL', 'BRA', 'CAN', 'CHE', 'CHN', 'DNK', 'DEU', 'ESP', 'FIN', 'FRA', 
                                   'GBR', 'IND', 'ISR', 'ITA', 'JPN', 'KOR', 'NLD', 'POL', 'RUS', 'SWE', 'USA', 'ZAF')
    )
    GROUP BY country, pubyear
    ORDER BY pub_cnt_frak DESC
    "),
    width = 100000, simplify=TRUE)
)
```
```{r, testing2, echo=FALSE, cache=TRUE}
# ISI approach
pubs_cntry_whole_wos_frac_ISI <- dbGetQuery(sql_pool, strwrap(paste0("
    SELECT country, pubyear, SUM(frac_share) pub_cnt_frak_ISI
    FROM(
      SELECT DISTINCT it.pubyear, inst.countrycode country, it.pk_items, frak.frac_share
      FROM wos_b_2018.items it
      JOIN wos_b_2018.items_authors_institutions iai ON it.pk_items=iai.fk_items
        AND (iai.role = 'author' OR iai.role IS NULL)
      JOIN wos_b_2018.institutions inst ON inst.pk_institutions=iai.fk_institutions
      JOIN wosb2018_frc_cntrylvl_org1 frak ON frak.fk_items = it.pk_items
        AND frak.countrycode = inst.countrycode
      WHERE it.pubyear = 2017
        AND it.doctype IN ('Article','Review')
        AND it.pubtype = 'Journal'
        AND inst.countrycode IN ('AUT', 'BEL', 'BRA', 'CAN', 'CHE', 'CHN', 'DNK', 'DEU', 'ESP', 'FIN', 'FRA', 
                                   'GBR', 'IND', 'ISR', 'ITA', 'JPN', 'KOR', 'NLD', 'POL', 'RUS', 'SWE', 'USA', 'ZAF')
    )
    GROUP BY country, pubyear
    ORDER BY pub_cnt_frak_ISI DESC
    "),
    width = 100000, simplify=TRUE)
)
```

```{r, echo=FALSE, cache=FALSE}
# table
pubs_cnt_compa <- cbind(pubs_cntry_whole_wos_frac, pubs_cntry_whole_wos_frac_ISI)
print(pubs_cnt_compa)
# kable(pubs_cnt_compa[1,3,4,6],
#       digits = 2
#       )

# kable(pubs_cntry_whole_wos_frac,
#       digits = 2
#       )
# kable(pubs_cntry_whole_wos_frac_ISI,
#       digits = 2
#       )
```

Only Italy and Korea swap places in this ranking of productivity.

# Scopus: scopus_b_2017

Die Datenqualität von Scopus erscheint signifikant besser zu sein, und zwar für den aktuellen Rand als auch insbesondere für vergangene Jahre.

```{r, echo=FALSE}
# Article und Review pro Jahr
papersByYear_sc <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT pk_items)
  FROM scopus_b_2017.items it
  WHERE pubtype = 'J'
  AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# inner vs left join between items and items_authors_institutions
outerInneDiff_sc <- dbGetQuery(sql_pool, strwrap(paste0("
SELECT tmp1.pubyear, (tmp2.count_outer - tmp1.count_inner) AS pub_diff
  FROM(
    SELECT pubyear, COUNT(DISTINCT iai.fk_items) count_inner
    FROM scopus_b_2017.items it
    JOIN scopus_b_2017.items_authors_institutions iai ON it.pk_items=iai.fk_items
    WHERE pubtype = 'J'
      AND doctype IN ('ar', 're')
      AND pubyear BETWEEN 2007 AND 2016
    GROUP BY pubyear) tmp1
  JOIN(
    SELECT pubyear, COUNT(DISTINCT iai.fk_items) count_outer
    FROM scopus_b_2017.items it
    LEFT JOIN scopus_b_2017.items_authors_institutions iai ON it.pk_items=iai.fk_items
    WHERE pubtype = 'J'
      AND doctype IN ('ar', 're')
      AND pubyear BETWEEN 2007 AND 2016
    GROUP BY pubyear) tmp2 ON tmp1.pubyear = tmp2.pubyear
  ORDER BY tmp1.pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# Article und Review mit fk_authors als NULL
paperNullAuthor_sc <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM scopus_b_2017.items it
  JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE fk_authors IS NULL
    AND pubtype = 'J'
    AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
    AND (type = 'RS' OR type IS NULL)
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# Article und Review mit fk_institutions als NULL
paperNullInstitution_sc <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM scopus_b_2017.items it
  JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE fk_institutions IS NULL
    AND pubtype = 'J'
    AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
    AND (type = 'RS' OR type IS NULL)
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# article or review with no information on type
paperNullType_sc <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM scopus_b_2017.items it
  JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  WHERE type IS NULL
    AND pubtype = 'J'
    AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
if(dim(paperNullType_sc)[1]==0){
  paperNullType_sc <- matrix(NA, nrow = length(2007:2016), ncol = 2)
  paperNullType_sc[,1] <- 2007:2016
  paperNullType_sc[,2] <- 0
}
```

```{r, echo=FALSE}
# article or review with no information on organization1
paperNullorganization1_sc <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(DISTINCT fk_items)
  FROM scopus_b_2017.items it
  JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
  WHERE organization1 IS NULL
    AND pubtype = 'J'
    AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
  GROUP BY pubyear
  ORDER BY pubyear ASC
  "),
  width = 100000, simplify = TRUE)
  )
```

```{r, echo=FALSE}
# all potential missing information jointly
paperOverall_sc <- dbGetQuery(sql_pool, strwrap(paste0("
  SELECT pubyear, COUNT(pk_items)
  FROM(
  SELECT DISTINCT pubyear, pk_items
  FROM scopus_b_2017.items it
  JOIN scopus_b_2017.items_authors_institutions iai ON it.pk_items = iai.fk_items
  WHERE pubtype = 'J'
    AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
    AND ((fk_institutions IS NULL
          AND (type = 'RS' OR type IS NULL)) OR
        (fk_authors IS NULL
          AND (type = 'RS' OR type IS NULL)) OR
        (type IS NULL))
  UNION
  SELECT DISTINCT pubyear,  pk_items
  FROM scopus_b_2017.items it
  JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
  JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
  WHERE organization1 IS NULL
    AND pubtype = 'J'
    AND doctype IN ('ar', 're')
    AND pubyear BETWEEN 2007 AND 2016
  )
  GROUP BY pubyear
  ORDER BY pubyear ASC  "),
  width = 100000, simplify = TRUE)
  )
```

Prozentualer Anteil von pk_items mit fehlenden Informationen (Scopus):

```{r, tab_friction_sc, echo=FALSE, cache=FALSE}
# print(outerInneDiff_sc[,2])
# print(paperNullType_sc[,2])
# print(paperNullInstitution_sc[,2])
# print(paperNullAuthor_sc[,2])
# print(paperNullorganization1_sc)
tab_friction_sc <- as.matrix(cbind(outerInneDiff_sc[,2]/papersByYear_sc[,2],
                                paperNullType_sc[,2]/papersByYear_sc[,2],
                                paperNullInstitution_sc[,2]/papersByYear_sc[,2],
                                paperNullAuthor_sc[,2]/papersByYear_sc[,2],
                                paperNullorganization1_sc[,2]/papersByYear_sc[,2],
                                (paperOverall_sc[,2] + outerInneDiff_sc[,2])/papersByYear_sc[,2])*100)
colnames(tab_friction_sc) <- c("no author nor affiliation",
                            "no author type",
                            "author w/o affiliation",
                            "affiliation w/o authors",
                            "no organization1",
                            "total")
rownames(tab_friction_sc) <- as.character(2007:2016)
kable(tab_friction_sc,
      digits = 1
      )
matplot(2007:2016, tab_friction_sc,
     xlim=c(2007,2016),
     ylim=c(0,100),
     lwd=2,
     lty = 1,
     bty="n",
     xlab="",
     ylab="%",
     type="l",
     col = rev(viridis(6)))
legend("topright", legend = rev(colnames(tab_friction)), fill=viridis(6), bty = "n")
```

<!-- ```{r, tab_dif_author_counting_sc, echo=FALSE, eval=FALSE} -->
<!-- # Differenz zwischen author_cnt und Zählung der fk_authors -->
<!-- dif_author_counting_sc <- dbGetQuery(sql_pool, strwrap(paste0(" -->
<!--   SELECT pubyear, SUM(diff) -->
<!--   FROM( -->
<!--     SELECT DISTINCT fk_items, pubyear, -->
<!--       CASE  WHEN author_cnt = (COUNT(DISTINCT fk_authors) OVER (PARTITION BY fk_items)) -->
<!--               THEN 0 -->
<!--             ELSE 1 -->
<!--       END diff -->
<!--     FROM scopus_b_2017.items it -->
<!--     JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items -->
<!--     WHERE pubtype = 'J' -->
<!--     AND doctype IN ('ar', 're') -->
<!--     AND pubyear BETWEEN 2007 AND 2016 -->
<!--     AND (type = 'RS' OR type IS NULL) -->
<!--   ) -->
<!--   GROUP BY pubyear -->
<!--   ORDER BY pubyear ASC -->
<!--   "), -->
<!--   width = 100000, simplify = TRUE) -->
<!--   ) -->
<!-- ``` -->

<!-- Prozentualer Anteil von pk_items mit Unterschieden in author_cnt und Zählung der zugehörigen fk_authors (Scopus): -->

<!-- ```{r, tab_diff_sc, echo=FALSE, eval=FALSE, cache=FALSE} -->
<!-- tmp <- as.matrix(dif_author_counting_sc[,2]/papersByYear_sc[,2] * 100) -->
<!-- rownames(tmp) <- as.character(2007:2016) -->
<!-- colnames(tmp) <- c("# Autoren != author_cnt") -->
<!-- kable(tmp, -->
<!--       digits = 1, -->
<!--       row.names = TRUE) -->
<!-- matplot(2007:2016, tmp, -->
<!--      xlim=c(2007,2016), -->
<!--      ylim=c(0,100), -->
<!--      lwd=2, -->
<!--      lty=1, -->
<!--      bty="n", -->
<!--      xlab="", -->
<!--      ylab="%", -->
<!--      type="l", -->
<!--      col=viridis(2)) -->
<!-- # legend("topright", -->
<!-- #        legend = c("# Autoren != author_cnt", "# Affiliation != inst_cnt"), -->
<!-- #        fill=viridis(2), -->
<!-- #        bty = "n") -->
<!-- ``` -->

## Implementation in Scopus

```sql
DROP TABLE items_iai_complete_scp;
DROP TABLE scopusb2017_frc_cntrylvl;

/*
Create temporaray table as an indicator of items with complete iai record:
*/
CREATE GLOBAL TEMPORARY TABLE items_iai_complete_scp (
    fk_items NUMBER
)
ON COMMIT DELETE ROWS;

/*
create empty table defining fractional weight for every paper by country
causes implicit commit
*/
CREATE TABLE scopusb2017_frc_cntrylvl (
    fk_items NUMBER,
    countrycode VARCHAR2(10 CHAR),
    frac_share NUMBER(3,2),
    ind_complete NUMBER(1,0)
);

/*
populate temporary table
*/
INSERT INTO items_iai_complete_scp (
SELECT DISTINCT tmp.fk_items -- subset of items with complete iai information
FROM( -- counts numbers of entries in iai for every item
    SELECT fk_items, COUNT (*) AS entry_cnt
    FROM scopus_b_2017.items it
    LEFT JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
    LEFT JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
    WHERE pubtype = 'J'
        AND doctype IN ('ar', 're')
        AND pubyear BETWEEN 2007 AND 2016
    GROUP BY fk_items
) tmp
JOIN( -- counts numbers of COMPLETE entries in iai for every item
    SELECT fk_items, COUNT (*) AS entry_cnt
    FROM scopus_b_2017.items it
    LEFT JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
    LEFT JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
    WHERE pubtype = 'J'
        AND doctype IN ('ar', 're')
        AND pubyear BETWEEN 2007 AND 2016
        AND fk_institutions IS NOT NULL
        AND fk_authors IS NOT NULL
        AND type IS NOT NULL
    GROUP BY fk_items
) tmp2 ON tmp.fk_items = tmp2.fk_items
AND tmp.entry_cnt = tmp2.entry_cnt); -- compare both numbers

/*
include items with COMPLETE iai information via fractional counting on author level:
*/
INSERT INTO scopusb2017_frc_cntrylvl
SELECT fk_items, countrycode, SUM(orga_frak) AS frac_share, 1 -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution of organsation on author level
      SELECT DISTINCT fk_items, ORGANIZATION1, countrycode, fk_authors,
        (1/(COUNT (DISTINCT fk_authors) OVER (PARTITION BY fk_items)))/(COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items, fk_authors)) AS orga_share
      FROM scopus_b_2017.items it
      JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
      JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE pubtype = 'J'
        AND doctype IN ('ar', 're')
        AND pubyear BETWEEN 2007 AND 2016
        AND type = 'RS'
        AND pk_items IN (SELECT fk_items FROM items_iai_complete_scp)
        AND organization1 IS NOT NULL
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode, 1;

/*
include items with INCOMPLETE iai information via fractional counting on organization level:
*/
INSERT INTO scopusb2017_frc_cntrylvl
SELECT fk_items, countrycode, SUM(orga_frak) AS frac_share, 0 -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution on organisation level
      SELECT DISTINCT fk_items, ORGANIZATION1, countrycode,
        (1/(COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items))) AS orga_share
      FROM scopus_b_2017.items it
      JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
      JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE pubtype = 'J'
        AND doctype IN ('ar', 're')
        AND pubyear BETWEEN 2007 AND 2016
        AND type = 'RS'
        AND pk_items NOT IN (SELECT fk_items FROM items_iai_complete_scp)
        AND organization1 IS NOT NULL
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode, 0;

CREATE INDEX scpb2018_frc_cntrylvl_ix ON scopusb2017_frc_cntrylvl (fk_items, countrycode);

COMMIT;
```
## Comparison

As an alternative implementation we also compute weights based on fractionation on the *organization1* level:

```sql

/*
create empty table defining fractional weight for every paper by country
based on fractionation on organization1 level
*/
CREATE TABLE scopusb2017_frc_cntrylvl_org1 (
    fk_items NUMBER,
    countrycode VARCHAR2(10 CHAR),
    frac_share NUMBER(3,2)
);

/*
include items with INCOMPLETE iai information via fractional counting on organization level:
*/
INSERT INTO scopusb2017_frc_cntrylvl_org1
SELECT fk_items, countrycode, SUM(orga_frak) AS frac_share -- aggregating on country level
FROM( -- aggregating on organization1 level
    SELECT fk_items, organization1, countrycode, SUM(orga_share)  AS orga_frak
    FROM( -- computing fractional contribution on organisation level
      SELECT DISTINCT fk_items, ORGANIZATION1, countrycode,
        (1/(COUNT (DISTINCT organization1) OVER (PARTITION BY fk_items))) AS orga_share
      FROM scopus_b_2017.items it
      JOIN scopus_b_2017.ITEMS_AUTHORS_INSTITUTIONS iai ON it.pk_items = iai.fk_items
      JOIN scopus_b_2017.institutions inst ON inst.pk_institutions = iai.fk_institutions
      WHERE pubtype = 'J'
        AND doctype IN ('ar', 're')
        AND pubyear BETWEEN 2007 AND 2016
        AND type = 'RS'
        AND organization1 IS NOT NULL
      )
    GROUP BY fk_items, organization1, countrycode
)
GROUP BY fk_items, countrycode;

CREATE INDEX scpb2017_frc_cntrylvl_org1_idx ON scopusb2017_frc_cntrylvl_org1 (fk_items, countrycode);

COMMIT;
```

Computing fractional counts of publications on national level for 2016:

```{r, testing_scp_1, echo=FALSE, cache=TRUE}
# hybrid approach
pubs_cntry_whole_wos_frac_scp <- dbGetQuery(sql_pool, strwrap(paste0("
    SELECT country, pubyear, SUM(frac_share) pub_cnt_frak
    FROM(
      SELECT DISTINCT it.pubyear, inst.countrycode country, it.pk_items, frak.frac_share
      FROM scopus_b_2017.items it
      JOIN scopus_b_2017.items_authors_institutions iai ON it.pk_items=iai.fk_items
      JOIN scopus_b_2017.institutions inst ON inst.pk_institutions=iai.fk_institutions
      JOIN scopusb2017_frc_cntrylvl frak ON frak.fk_items = it.pk_items
        AND frak.countrycode = inst.countrycode
      WHERE it.pubyear = 2016
        AND it.doctype IN ('ar','re')
        AND it.pubtype = 'J'
        AND inst.countrycode IN ('AUT', 'BEL', 'BRA', 'CAN', 'CHE', 'CHN', 'DNK', 'DEU', 'ESP', 'FIN', 'FRA',
                                   'GBR', 'IND', 'ISR', 'ITA', 'JPN', 'KOR', 'NLD', 'POL', 'RUS', 'SWE', 'USA', 'ZAF')
    )
    GROUP BY country, pubyear
    ORDER BY pub_cnt_frak DESC
    "),
    width = 100000, simplify=TRUE)
)
```

```{r, testing_scp_2, echo=FALSE, cache=TRUE}
# ISI approach
pubs_cntry_whole_wos_frac_ISI_scp <- dbGetQuery(sql_pool, strwrap(paste0("
    SELECT country, pubyear, SUM(frac_share) pub_cnt_frak_ISI
    FROM(
      SELECT DISTINCT it.pubyear, inst.countrycode country, it.pk_items, frak.frac_share
      FROM scopus_b_2017.items it
      JOIN scopus_b_2017.items_authors_institutions iai ON it.pk_items=iai.fk_items
      JOIN scopus_b_2017.institutions inst ON inst.pk_institutions=iai.fk_institutions
      JOIN scopusb2017_frc_cntrylvl_org1 frak ON frak.fk_items = it.pk_items
        AND frak.countrycode = inst.countrycode
      WHERE it.pubyear = 2016
        AND it.doctype IN ('ar','re')
        AND it.pubtype = 'J'
        AND inst.countrycode IN ('AUT', 'BEL', 'BRA', 'CAN', 'CHE', 'CHN', 'DNK', 'DEU', 'ESP', 'FIN', 'FRA',
                                   'GBR', 'IND', 'ISR', 'ITA', 'JPN', 'KOR', 'NLD', 'POL', 'RUS', 'SWE', 'USA', 'ZAF')
    )
    GROUP BY country, pubyear
    ORDER BY pub_cnt_frak_ISI DESC
    "),
    width = 100000, simplify=TRUE)
)
```

```{r, echo=FALSE, cache=FALSE}
# table
pubs_cnt_compa <- cbind(pubs_cntry_whole_wos_frac_scp, pubs_cntry_whole_wos_frac_ISI_scp)
print(pubs_cnt_compa)
# kable(pubs_cnt_compa[1,3,4,6],
#       digits = 2
#       )

# kable(pubs_cntry_whole_wos_frac,
#       digits = 2
#       )
# kable(pubs_cntry_whole_wos_frac_ISI,
#       digits = 2
#       )
```

# Summing up

The implications of the adapted counting method seems rather minor on the chosen analysis level of countries.

Fractionation on the author level relies on complete information on the affiliations of single authors.

In the WoS the share of items with incomplete records constantly constitutes a substantial part of the corpus. Although data quality has strongly improved since 2008, still for 2017 more than 10% of all articles and reviews are incomplete. Furthermore data quality apparently gets improved every year, but the rate of improvement has slowed down considerably in the last years.

Socpus holds less incomplete recods, their share has been below 10% for all years since the first observation year 2007.

Consequently fractionizing on the author level seems only reasonable on complete records, whereas incomplete records might be fractionoized on the *organization1* level. Both approaches could be improved via a wordwide institution harmonization, which has partially be done by ISI Karlsruhe for their EFI reports.


```{r, teardown, include=FALSE}
poolClose(sql_pool)
```


# Literatur
